\documentclass{llncs}
\usepackage{times}
\usepackage{subcaption}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx,wrapfig,lipsum}
%\usepackage{tikz}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{adjustbox}
%\usepackage{ulem}
\usepackage[colorlinks,linkcolor=black,anchorcolor=black,
    citecolor=blue,urlcolor=black,bookmarks=true]{hyperref}
\usepackage{pifont}

\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\newcommand{\code}[1]{{\small {\ensuremath{\tt #1}}}}
\newcommand\LJ[1]{\textcolor{green}{#1}}
\newcommand\LL[1]{\textcolor{red}{#1}}
\newcommand\SJ[1]{\textcolor{blue}{#1}}
\newcommand\loc[1]{\textcolor{blue}{To do: #1}}

%\newtheorem{definition}{Definition}
%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{Lemma}
%\newtheorem{example}{Example}
%\newtheorem{proposition}{Proposition}
\setcounter{tocdepth}{3}
\pagestyle{plain}

\begin{document}

%\setcopyright{acmcopyright}

\title{Loop-invariant Generation through Active Learning}
% \author{
% Jiaying Li, Li Li, Le Guang Loc, Jun Sun\\
% \institute{
% Singapore University of Technology and Design \\
% \email{\{jiaying\_li,li\_li,quangloc\_le,sunjun\}@sutd.edu.sg}}
% }

\author{Jiaying Li\inst{1}, Jun Sun\inst{1}, Li Li\inst{1}, Quang Loc Le\inst{1} and Shang-Wei Lin\inst{2}}
\institute{Singapore University of Technology and Design
\and
School of Computer Science and Engineering, Nanyang Technological University
}

%%\numberofauthors{4}
%\author{
%% You can go ahead and credit any number of authors here,
%% e.g. one 'row of three' or two rows (consisting of one row of three
%% and a second row of one, two or three).
%%
%% The command \alignauthor (no curly braces needed) should
%% precede each author name, affiliation/snail-mail address and
%% e-mail address. Additionally, tag each line of
%% affiliation/address with \affaddr, and tag the
%% e-mail address with \email.
%%
%% 1st. author
%\alignauthor
%Jiaying Li\\
%       \affaddr{Singapore University of Technology and Design}\\
%       \email{jiaying\_li@mymail.sutd.edu.sg}
%% 2nd. author
%\alignauthor
%Li Li\\
%       \affaddr{Singapore University of Technology and Design}\\
%       \email{li\_li@sutd.edu.sg}
%% 3rd. author
%\and  % use '\and' if you need 'another row' of author names
%\alignauthor
%Le Guang Loc\\
%       \affaddr{Singapore University of Technology and Design}\\
%       \email{guangloc\_le@sutd.edu.sg}
%% 4th. author
%\alignauthor
%Jun Sun\\
%       \affaddr{Singapore University of Technology and Design}\\
%       \email{sunjun@sutd.edu.sg}
%}

\maketitle

\begin{abstract}
Loop invariant generation is important in program analysis and verification. In this work, we propose a technique for automatic loop-invariant generation through a combination of active learning and verification. Given a Hoare triple of a program containing a loop, we start with randomly testing the program. We collect program states at run-time and categorize them based on whether they satisfy the invariant to be discovered. Next, classification techniques are employed to generate candidate loop invariants. In particular, we refine the candidates through active learning so as to overcome the lack of sampled program states.
Only after the candidate invariant cannot be improved further through active learning, we verify whether a candidate can be used to prove the Hoare triple.
If it cannot, the generated counterexample is used for re-classification and we repeat the above process. Furthermore, we show that by introducing
path-sensitive learning, i.e., partitioning the program states according to program locations they visit and classifying each partition separately, we are able to learn disjunctive loop invariants. We have developed a prototype tool and applied it to verify a set of benchmark programs. The evaluation shows that our approach complements existing approaches.

%-------------------------------------------------------------------------
%  \keywords Loop Invariant $\cdot$ Active Learning $\cdot$ Verification
% $\cdot$ Program Analysis
%-------------------------------------------------------------------------
\end{abstract}

\input{introduction.tex}
\input{datacollection.tex}
\input{classifierlearning.tex}
%\input{activelearning.tex}
\vspace{-0.3cm}
\input{evaluation.tex}
\input{related.tex}

\newpage

\bibliographystyle{abbrv}
\bibliography{zeno}

\end{document}
